### 2020-09-11 Paper/results.tex
folder name

\subsection{Extension Methods for Comparison}
To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}. We also evaluate the performance of reference forecasting algorithms that could be used for extending such signals. These methods are:
\begin{itemize}
\item The EDMD has been developed by Williams \etal~\cite{Williams15data}. The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows catching dynamic of nonlinear systems~\cite{Korda18linear}.
\item The GPR~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model. That one is based on the Gaussian process structure, and therefore offers more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
\item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, seasonal and ARMA components, with a specific dynamic for the seasonal component. This model demands the estimation of numerous parameters and, by implication, may be slow. 
\end{itemize}

@@

\paragraph{Influence of the Noise Variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$. Then, the forecasting algorithm is run on $500$ realizations on the discrete signal $\bx$ for $200$ different values of $\sigma$, logarithmically equispaced from $10^{-7}$ to $10^{-1}$. For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$. Fig.~\ref{fig:res.noise.sine} shows the the experimental forecasting variance in function of the noise variance for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.

@@

\paragraph{Influence of the Training Dataset Size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$. Then, {\sf SigExt} is run on $500$ realizations of the discrete signal $\bx$ for $200$ different values of training dataset size $K$, logarithmically equispaced from $450$ to $2000$. For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$. Fig.~\ref{fig:res.size.sine} shows the the experimental forecasting variance in function of the training dataset size for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.

@@

As anticipated by Theorem~\ref{th:error}, the variance of the estimator decreases linearly with $\frac1K$ as long as the dataset size is large enough. Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance in function of $K$ tends towards an asymptote of slope $-1$.

@@

From that long signal, we build a dataset of $379$ non\-overlapping signals of 60 seconds, \ie~$N=6000$. For each segment, we apply the forecasting methods introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}. We forecast $7$-second-long extensions on each segment of the signal, corresponding to $L =700$. Thus, in order to catch the dynamical behavior, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$. As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$. The extensions obtained on one of these subsignals are shown in Fig.~\ref{fig:tho}. The resulting MSEs of different methods are given in Table~\ref{tab:THO}. Note that the results are given in ``mean $\pm$ standard deviation'' format. The TBATS extension method is not implemented, as its excessive computing time makes it impossible to implement it on $6000$ segments within a reasonable period of time.
%
Note that the MSE of {\sf SigExt} is in average higher than the MSE of EDMD or the symmetric extension, with a huge standard deviation. This variability is caused by the presence of few segments contaminated by artifacts so that it is unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}. The left of Fig.~\ref{fig:THO.failure} illustrates one of these outliers, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal. The EDMD and symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}. Nevertheless, {\sf SigExt} provides a sufficiently relevant extension to give TF representations sparingly affected by boundary effects. On the right of Fig.~\ref{fig:THO.failure}, we display a comparison between the right boundary of SST of the same segment of signal (top-right), and its boundary-free SST obtained after the {\sf SigExt} forecasting (bottom-right). The extension of the instantaneous frequency visible on the right side of the image illustrates the reduction of boundary effects produced despite an inaccurate signal forecasting.

@@

We set the extension length $L$ accordingly to the window length used by the TF analysis tool. For instance, the window length used to evaluate the STFT is of $1400$ samples. To prevent the STFT from being sensitive to the boundaries, we set $L=700$. In this way, the evaluation of the spectral content of the signal near its boundary is not limited by a lack of information all along the window support. From now on, all results are given for $L$ equal to the half of the width of the window used in the TF transform.
%
In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf} by evaluating the whole TF representations (including boundaries). Even though {\sf SigExt} performs moderately well in the sense of MSE, the boundary effects are dramatically reduced on the TF representations, and the averaged performance is in the same order of that given by EDMD or GPR. A t-test is performed to compare the performance index of {\sf SigExt} with those of other methods. Under the null hypothesis that the means are equal, with a $5\%$ significance level, the t-tests show no statistical significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.

@@

We divide the signal into 32-second-long segments and apply {\sf BoundEffRed} to each of them. Table~\ref{tab:otd.ppg} shows the performance index $D$ of the boundary-free TF representations, averaged over the signals. For all the TF representations considered, the results show that {\sf SigExt} reduces boundary effects about as efficiently as the extensions given by EDMD or GPR. As for the respiratory signal, t-tests show no statistical significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered. For a visual inspection, the TF representation of SST resulting from the {\sf BoundEffRed} strategy is shown in the bottom-right panel of Fig.~\ref{fig:ex.intro}, where SST is applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}. It produces a significant improvement in the quality of the SST near boundaries. Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished. Real-time tracking of the instantaneous frequencies contained in the measured signal is therefore greatly facilitated.

@@

To further evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add Gaussian noise to the measured PPG signal. It is thus an additional noise to the measurement noise actually contained in the signal. Fig.~\ref{fig:otd.noise} shows the averaged performance index of {\sf BoundEffRed} for different values of the Signal-to-Noise Ratio (SNR). Whatever the representation considered, {\sf BoundEffRed} is relatively robust to noise, as long as the SNR remains above $10$~dB. Below this level, the reduction of boundary effects gradually deteriorates.

>>>>

\subsection{Extension Methods for Comparison}
To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}. We also evaluate the performance of reference forecasting algorithms that could be used for extending such signals. These methods are:
\begin{itemize}
\item The EDMD has been developed by Williams \etal~\cite{Williams15data}. The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows catching dynamic of nonlinear systems~\cite{Korda18linear}.
\item The GPR~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model. That one is based on the Gaussian process structure, and therefore offers more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
\item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, seasonal and ARMA components, with a specific dynamic for the seasonal component. This model demands the estimation of numerous parameters and, by implication, may require a long computation time.
\end{itemize}

@@

\paragraph{Influence of the Noise Variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$. Then, the forecasting algorithm is run on $500$ realizations on the discrete signal $\bx$ for $200$ different values of $\sigma$, logarithmically equispaced from $10^{-7}$ to $10^{-1}$. For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$. Fig.~\ref{fig:res.noise.sine} shows the experimental forecasting variance as a function of the noise variance for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.

@@

\paragraph{Influence of the Training Dataset Size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$. Then, {\sf SigExt} is run on $500$ realizations of the discrete signal $\bx$ for $200$ different values of training dataset size $K$, logarithmically equispaced from $450$ to $2000$. For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$. Fig.~\ref{fig:res.size.sine} shows the experimental forecasting variance as a function of the training dataset size for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.

@@

As anticipated by Theorem~\ref{th:error}, the variance of the estimator decreases linearly with $\frac1K$ as long as the dataset size is large enough. Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance as a function of $K$ tends towards an asymptote of slope $-1$.

@@

From that long signal, we build a dataset of $379$ non\-overlapping signals of 60 seconds, \ie~$N=6000$. For each segment, we apply the forecasting methods introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}. We forecast $7$-second-long extensions on each segment of the signal, corresponding to $L =700$. Thus, in order to catch the dynamical behavior, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$. As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$. The extensions obtained on one of these subsignals are shown in Fig.~\ref{fig:tho}. The resulting MSEs of different methods are given in Table~\ref{tab:THO}. Note that the results are given in ``mean $\pm$ standard deviation'' format. The TBATS extension method is not implemented, as its excessive computing time makes it impossible to implement it on $6000$ segments within a reasonable period of time.
%
Note that the MSE of {\sf SigExt} is on average higher than the MSE of EDMD or the symmetric extension, with a huge standard deviation. This variability is caused by the presence of few segments contaminated by artifacts so that it is unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}. The left of Fig.~\ref{fig:THO.failure} illustrates one of these outliers, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal. The EDMD and symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}. Nevertheless, {\sf SigExt} provides a sufficiently relevant extension to give TF representations sparingly affected by boundary effects. On the right of Fig.~\ref{fig:THO.failure}, we display a comparison between the right boundary of SST of the same segment of signal (top-right), and its boundary-free SST obtained after the {\sf SigExt} forecasting (bottom-right). The extension of the instantaneous frequency visible on the right side of the image illustrates the reduction of boundary effects produced despite an inaccurate signal forecasting.

@@

We set the extension length $L$ accordingly to the window length used by the TF analysis tool. For instance, the window length used to evaluate the STFT is of $1400$ samples. To prevent the STFT from being sensitive to the boundaries, we set $L=700$. In this way, the evaluation of the spectral content of the signal near its boundary is not limited by a lack of information in the interval delimited by the window support. From now on, all results are given for $L$ equal to the half of the width of the window used in the TF transform.
%
In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf} by evaluating the whole TF representations (including boundaries). Even though {\sf SigExt} performs moderately well in the sense of MSE, the boundary effects are dramatically reduced on the TF representations, and the averaged performance is in the same order of that given by EDMD or GPR. A t-test is performed to compare the performance index of {\sf SigExt} with those of other methods. Under the null hypothesis that the means are equal, with a $5\%$ significance level, the t-tests show no statistically significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.

@@

We divide the signal into 32-second-long segments and apply {\sf BoundEffRed} to each of them. Table~\ref{tab:otd.ppg} shows the performance index $D$ of the boundary-free TF representations, averaged over the signals. For all the TF representations considered, the results show that {\sf SigExt} reduces boundary effects about as efficiently as the extensions given by EDMD or GPR. As for the respiratory signal, t-tests show no statistically significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered. For visual inspection, the TF representation of SST resulting from the {\sf BoundEffRed} strategy is shown in the bottom-right panel of Fig.~\ref{fig:ex.intro}, where SST is applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}. It produces a significant improvement in the quality of the SST near boundaries. Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished. Real-time tracking of the instantaneous frequencies contained in the measured signal is therefore greatly facilitated.

@@

To further evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add Gaussian noise to the measured PPG signal. It is thus an additional noise to the measurement noise actually contained in the signal. Fig.~\ref{fig:otd.noise} shows the average performance index of {\sf BoundEffRed} for different values of the Signal-to-Noise Ratio (SNR). Whatever the representation considered, {\sf BoundEffRed} is relatively robust to noise, as long as the SNR remains above $10$~dB. Below this level, the reduction of boundary effects gradually deteriorates.
